---
title: "CVXR for PortfolioAnalytics"
author: "Xinran Zhao"
date: "2023/02/14"
output: rmarkdown::pdf_document
bibliography: PA.bib
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{CVXR_in_PA_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
Sys.setlocale("LC_TIME", "English")
```

\tableofcontents
\newpage
# 1 Introduction
CVXR is an R package that provides an object-oriented modeling language for convex optimization, including the Second-Order Cone Optimization(SOCopt) required to minimize Expected Quadratic Shortfall(EQS) problem, which is not supported by other solvers in PortfolioAnalytics. Hence, CVXR is a great extension of PortfolioAnalytics.

The purpose of this vignette is to demonstrate examples of optimization problems that can be solved in PortfolioAnalytics with CVXR and its many supported solvers. The problem types covered include not only Linear Programming(LP), Quadratic Programming(QP) but also Second-Order Cone Programming(SOCP). Multiple solvers supported by CVXR can be selected according to optimization types. For example, SCS and ECOS can completely cover the types of problems that ROI can deal with, such as mean-variance and ES problem. In order to better understand the functionality and use of PortfolioAnalytics, users are recommended to read the Vignette [*Introduction to PortfolioAnalytics*](https://cran.r-project.org/web/packages/PortfolioAnalytics/vignettes/portfolio_vignette.pdf) first.

# 2 Getting Started
## 2.1 Load Packages
Load the necessary packages.
```{r, message=FALSE}
library(PortfolioAnalytics)
library(CVXR)
library(data.table)
library(xts)
library(PCRA)
```

## 2.2 Solvers
The website https://cvxr.rbind.io/ shows that CVXR currently supports the use of 9 solvers, some of which are commercial (CBC, CPLEX, GUROBI, MOSEK)^[MOSEK ,GUROBI, and CPLEX require licenses to use them but free academic licenses are available for all three.] and the others are open source(GLPK, GLPK_MI, OSQP, SCS, ECOS). 

The PortfolioAnalytics package provides following two main wrapper functions for constrainted optimization of portfolios by a wide variety of methods:

`optimize.portfolio(R =, portfolio =, optimize_method =)`

`optimize.portfolio.rebalancing(R =, portfolio =, optimize_method =, rebalance_on =, training_period =, rolling_window =)`

Different solvers support different types of portfolio optimization problems and should be specified by the argument `optimize_method`. The `optimize_method=c("CVXR", {CVXRsolver})` argument of the function `optimize.portfolio` and `optimize.portfolio.rebalancing` allows the user to specify the solver to use with CVXR. If the argument is `optimize_method="CVXR"`, the default solver for LP and QP type portfolio optimization problems such as maximum mean return and minimum variance portfolio optimization is OSQP, and the default solver for SOCP type portfolio optimizations, such as "robust portfolio optimization" to control for alpha uncertainty, and Expected Quadratic Shortfall (EQS) portfolio optimization, is ECOS.

| Solver | LP | QP | SOCP |
| :-- | :----: | :----: | :----: |
|CBC|$\checkmark$| | |
|GLPK|$\checkmark$| | |
|GLPK_MI|$\checkmark$| | |
|OSQP|$\colorbox{gray}{\checkmark}$|$\colorbox{gray}{\checkmark}$| |
|SCS|$\checkmark$|$\checkmark$|$\checkmark$|
|ECOS|$\checkmark$|$\checkmark$|$\colorbox{gray}{\checkmark}$|
|CPLEX|$\checkmark$|$\checkmark$|$\checkmark$|
|GUROBI|$\checkmark$|$\checkmark$|$\checkmark$|
|MOSEK|$\checkmark$|$\checkmark$|$\checkmark$|

## 2.3 Data
The edhec data set from the PerformanceAnalytics package is used as example data for examples in Section 3 to Section 8. The edhec data set is an xts object that contains monthly returns for 13 assets from 1997-01 to 2019-11. We use the edhec data of the last 5 years as the example data to show how to use the code.
```{r}
data(edhec)
# Use edhec for a returns object
ret_edhec <- tail(edhec, 60)
colnames(ret_edhec) <- c("CA", "CTAG", "DS", "EM", "EMN", "ED", "FIA", 
                       "GM", "LSE", "MA", "RV", "SS", "FF")
print(head(ret_edhec, 5))
# Get a character vector of the asset names
fund_edhec <- colnames(ret_edhec)
```

`tsPlotMP` is a function of R package `PCRA` which can plot time series for the return data.
```{r}
tsPlotMP(ret_edhec, layout = c(2, 7))
```
\begin{center}
Fig 2.1
\end{center}
The CRSP data set is an xts object that contains the daily returns of stocks from 1993-01 to 2015-12^[CRSP: Center for Research in Security Prices, LLC.]. We select the top 30 small cap stocks based on the market capitalization, and use this larger and more frequent data set to show more meaningful and interesting results in Section 9. We don't want to use the large data set everywhere to slow down the code or distract the main point.

```{r}
# use CRSP daily data set
stocks <- stocksCRSPdaily
returnMat <- tapply(stocks[, Return], list(stocks$Date, stocks$TickerLast), I)
stocksCRSP_D <- xts(returnMat, as.Date(rownames(returnMat)))

# top 30 small cap stocks based on the market capitalization
small_cap <- factorsSPGMI[CapGroupLast == "SmallCap"]
sc_cap <- small_cap[,mean(LogMktCap),by="TickerLast"]
sc_cap <- sc_cap[order(sc_cap$V1, decreasing = TRUE),]
sc_30 <- sc_cap[,TickerLast][1:30]
ret_CRSP <- stocksCRSP_D[, sc_30]

print(head(ret_CRSP, 3))
fund_CRSP <- colnames(ret_CRSP)
```

In the following part, we only show the time series of monthly returns of 10 CRSP stocks in the last five years, but you can use this code to check the time series performance of all stocks in any frequency and any time period.
```{r}
# generate monthly return in last 5 years
ep <- endpoints(ret_CRSP, on= "months", k=1)
prod1 <- function(x){apply(x+1, 2, prod)}
retM_CRSP <- period.apply(ret_CRSP, INDEX = ep, FUN = prod1) - 1
retM_CRSP_5 <- tail(retM_CRSP, 60)

# time series plot of 10 stocks
tsPlotMP(retM_CRSP_5[, 1:10])
```
\begin{center}
Fig 2.2
\end{center}
## 2.4 Optimization Problems
In this Vignette, all mean vectors and covariance matrices in the optimization formula will use standard sample based estimates. All optimization problems treated will use linear constraints unless stated otherwise. There will be one equality constraint, i.e., the full-investment constraint, and one or more inequality constraints such as the long-only and box constraints. More comprehensive constraint types can be found in the vignette Ross (2018) [*Introduction to PortfolioAnalytics*](https://cran.r-project.org/web/packages/PortfolioAnalytics/vignettes/portfolio_vignette.pdf). 

This vignette will be organized by objective type and provide some visual examples.

# 3 Maximizing Mean Return
The objective to maximize mean return is a linear problem of the form:
$$\max\limits_{w} \quad \boldsymbol{\mu}'\boldsymbol{w}$$
\begin{equation*}
\begin{aligned}
s.t. \quad & A\boldsymbol{w} \ge b\\
& B\boldsymbol{w} = c
\end{aligned}
\end{equation*}

Where $\boldsymbol{\mu}$ is the estimated asset returns mean vector and $\boldsymbol{w}$ is the vector of portfolio weights.

## 3.1 Portfolio Object
The first step in setting up a model is to create the portfolio object. Portfolio object should be created in the form of constraints and objective specifications. In the following we create full-investment and box constraints specifications, and a maximum return objective specification.
```{r}
# Create portfolio object
pspec_maxret <- portfolio.spec(assets=fund_edhec)
# Add constraints to the portfolio object
pspec_maxret <- add.constraint(pspec_maxret, type="full_investment")
pspec_maxret <- add.constraint(portfolio = pspec_maxret, type = "box",
                              min = rep(0.02, 13),
                              max = c(rep(0.6, 4), rep(0.2, 9)))
# Add objective to the portfolio object
pspec_maxret <- add.objective(portfolio = pspec_maxret, 
                             type = "return", name = "mean")
pspec_maxret
```

## 3.2 Optimization
The next step is to run the optimization. Note that `optimize_method=c("CVXR", {CVXRsolver})` should be specified in the function `optimize.portfolio` to use CVXR solvers for the optimization, or use the default solver by giving `optimize_method="CVXR"`. For maximizing mean return problem, which is a linear programming, the default solver is `OSQP`.
```{r}
# Run the optimization with default solver
opt_maxret <- optimize.portfolio(R=ret_edhec, portfolio=pspec_maxret, 
                                 optimize_method="CVXR", trace=TRUE)
opt_maxret
opt_maxret$solver

# Run the optimization with specific solver
opt_maxret_glpk <- optimize.portfolio(R=ret_edhec, portfolio=pspec_maxret, 
                                 optimize_method=c("CVXR", "GLPK"), trace=TRUE)
opt_maxret_glpk$solver
```

## 3.3 Backtesting
An out of sample backtest is run with `optimize.portfolio.rebalancing`. In this example, an initial training period of 36 months is used and the portfolio is rebalanced quarterly.
```{r warning=FALSE}
bt_maxret <- optimize.portfolio.rebalancing(R=ret_edhec, portfolio=pspec_maxret,
                                            optimize_method="CVXR",
                                            rebalance_on="quarters",
                                            training_period=36)
```


The call to `optimize.portfolio.rebalancing` returns the `bt_maxret` object which is a list containing the optimal weights and objective measure at each rebalance period.

```{r}
class(bt_maxret)
names(bt_maxret)
```

# 4 Minimizing Variance
The objective to minimize variance is a quadratic problem of the form:
$$\min\limits_{w} \quad \boldsymbol{w}'\Sigma \boldsymbol{w}$$
subject to only the full-investment constraint, where $\Sigma$ is the estimated covariance matrix of asset returns and $\boldsymbol{w}$ is the set of weights. It is a quadratic problem.

## 4.1 Global Minimum Variance Portfolio
### 4.1.1 Portfolio Object
In this example, the only constraint specified is the full investment constraint, therefore the optimization problem is solving for the global minimum variance portfolio.
```{r}
# Create portfolio object
pspec_gmv <- portfolio.spec(assets=fund_edhec)
# Add full-investment constraint
pspec_gmv <- add.constraint(pspec_gmv, type="full_investment")
# Add objective of minimizing variance
pspec_gmv <- add.objective(portfolio = pspec_gmv, type = "risk", name = "var")
```

### 4.1.2 Optimization
```{r}
opt_gmv <- optimize.portfolio(ret_edhec, pspec_gmv, optimize_method = "CVXR")
opt_gmv
```
As this example illustrates, a global minimum variance portfolio can have short positions.

## 4.2 Lone-Only and Group Constrained Minimum Variance Portfolio 
Various linear inequality constraint, such as box constraints, group constraints and a target mean return constraint, can be used with GMV portfolio construction. Here we demonstrate the case of linearly constrained minimum variance portfolio.

```{r}
# portfolio object
pspec_mv <- add.constraint(pspec_gmv, type = "long_only")
pspec_mv <- add.constraint(pspec_mv, type = "group",
                            groups=list(groupA=1,
                                        groupB=c(2:12),
                                        groupC=13),
                            group_min=c(0, 0.05, 0.05),
                            group_max=c(0.4, 0.8, 0.5))
pspec_mv <- add.constraint(pspec_mv, type = "return", return_target=0.003)
pspec_mv

# optimization
opt_mv <- optimize.portfolio(ret_edhec, pspec_mv, optimize_method = "CVXR")
opt_mv

# backtesting
bt_mv <- optimize.portfolio.rebalancing(R=ret_edhec, portfolio=pspec_mv,
                                            optimize_method="CVXR",
                                            rebalance_on="quarters",
                                            training_period=36)
```
The use of an alternative to the CVXR default solver will get the same result to many significant digits. In this example we use `optimize_method=c("CVXR", "ECOS")`, since `OSQP` is the default solver, and get the very similar results.
```{r}
opt_mv_ecos <- optimize.portfolio(ret_edhec, pspec_mv, optimize_method = c("CVXR", "ECOS"))
opt_mv_ecos

opt_mv$solver
opt_mv_ecos$solver
```

# 5 Maximizing Quadratic Utility
Next we demonstrate the classical quadratic utility form of Markowitz's mean-variance model, where the quadratic utility function is $\rm QU(\boldsymbol{w}) = \mu_p - \lambda \sigma_p^2 = \boldsymbol{\mu'w}-\lambda \boldsymbol{w'}\Sigma \boldsymbol{w}$:
$$\max\limits_{w} \quad \boldsymbol{\mu}'\boldsymbol{w} - \lambda\boldsymbol{w}'\Sigma\boldsymbol{w}$$
$$s.t. \quad A\boldsymbol{w} \ge b$$
Where $\boldsymbol{\mu}$ is the estimated mean asset returns, $0 \le \lambda < \inf$ is the risk aversion parameter, $\Sigma$ is the estimated covariance matrix of asset returns and $\boldsymbol{w}$ is the set of weights. Quadratic utility maximizes return while penalizing variance. The risk aversion parameter $\lambda$ controls how much portfolio variance is penalized, and when $\lambda=0$ it becomes a maximum mean return problem of Section 3, and as $\lambda \rightarrow \inf$, it becomes the minimum variance problem of Section 4.

## 5.1 Portfolio Object
In this case the objectives of the portfolio should be both return and risk, and for this example we will use a risk aversion parameter $\lambda$ to be 20 by setting `risk_aversion = 20`.
```{r}
pspec_mvo <- portfolio.spec(assets=fund_edhec)
pspec_mvo <- add.constraint(pspec_mvo, type="full_investment")
pspec_mvo <- add.constraint(pspec_mvo, type="long_only")
# Add objectives
pspec_mvo <- add.objective(portfolio = pspec_mvo, type = "return", name = "mean")
pspec_mvo <- add.objective(portfolio = pspec_mvo, type = "risk", name = "var",
                           risk_aversion = 20)
```

## 5.2 Optimization
The optimization result `opt_mvo` shows the call, optimal weights, and the objective measure. Objective measure contains quadratic utility, mean return and standard deviation.
```{r}
opt_mvo <- optimize.portfolio(ret_edhec, pspec_mvo, optimize_method = "CVXR")
opt_mvo
```

# 6 Minimizing Expected Shortfall
Expected Shortfall(ES) is also called Conditional Value-at-Risk(CVaR) and Expected Tail Loss(ETL). The ES of a portfolio is
\begin{equation*}
\begin{aligned}
ES_\gamma(r_P) 
= ES_\gamma(\boldsymbol{w}) 
&= -E(r_P|r_P \le q_\gamma(\boldsymbol{w}))\\
&=-E(\boldsymbol{w'r}|\boldsymbol{w'r} \le q_\gamma(\boldsymbol{w}))
\end{aligned}
\end{equation*}

where $r_P$ is a random return of a portfolio $P$, and $\boldsymbol{r}$ is the loss return which is negative, and $q_\gamma$ is $\gamma$-quantile and $\gamma$ is usually a "tail" probability such as 0.01, 0.05, in which case ES is a tail risk measure. But one could also choose $\gamma=0.25$ or $\gamma=0.5$, in which case ES is just a "downside" risk measure, and if $\gamma>0.5$, the problem will take $1-\gamma$ as the tail probability.

It was shown by @rockafellar2000optimization that the optimal minimum ES portfolio is the result of the minimization:
$$\min \limits_{\boldsymbol{w}} ES_\gamma(\boldsymbol{w}) = \min \limits_{\boldsymbol{w}, t} F_\gamma(\boldsymbol{w}, t)$$
where
$$F_\gamma(\boldsymbol{w},t)=-t+\frac{1}{\gamma} \int[t-\boldsymbol{w'r}]^+ \cdot f(\boldsymbol{r})d\boldsymbol{r}$$
by replacing $q_\gamma$ with the free variable $t$, and with the discrete data the formula is:
$$\hat{F}_\gamma(\boldsymbol{w}, t) = -t + \frac{1}{n \cdot \gamma} \sum_{i=1}^n[t-\boldsymbol{w'r_i}]^+$$
The positive part function, $[t-\boldsymbol{w'r_i}]^+$, can easily be converted to a collection of linear constraints, hence, the minimization of ES is equivalent to solving a linear programming problem.

The ES objective is in the form of:
$$\min\limits_{\boldsymbol{w}, t} \quad -t + \gamma^{-1}E(t-\boldsymbol{w'r_i})^+$$
where $0<\gamma<1$ is the tail probability value, and $t$ is the value from which shortfalls are measured in the optimal solution. Many authors also use $p$ or $\alpha$ as the quantile, e.g., in @rockafellar2000optimization and other vignettes of PortfolioAnalytics, and use $\eta$ as the risk measure variable, e.g., in @krokhmal2007higher.

## 6.1 Portfolio Object
The default probability is $\gamma = 5\%$. Specific probability could be given by `arguments`.
```{r}
pspec_es <- portfolio.spec(assets=fund_edhec)
pspec_es <- add.constraint(pspec_es, type="full_investment")
# Add objective of minimizing ES by using the default gamma
pspec_es <- add.objective(portfolio = pspec_es, type = "risk", name = "ES")
# Add objective of minimizing ES by using the specific gamma=0.1
pspec_es_1 <- add.objective(portfolio = pspec_es, type = "risk", name = "ES",
                          arguments = list(p=0.1))
```

## 6.2 Optimization
```{r}
# GMES with default gamma=0.05
opt_es <- optimize.portfolio(ret_edhec, pspec_es, optimize_method = "CVXR")
opt_es
# GMES with specific gamma=0.1
opt_es_1 <- optimize.portfolio(ret_edhec, pspec_es_1, optimize_method = "CVXR")
opt_es_1
```

# 7 Minimizing Expected Quadratic Shortfall
Expected Quadratic Shortfall(EQS) is also called Second-Moment Coherent Risk Measure(SMCR). The objective to minimize EQS is in the form of:
$$\min\limits_{\boldsymbol{w}, t} \quad -t + \gamma^{-1}||(t-\boldsymbol{w'r_i})^+||_2$$
where $\gamma$ is the tail probability and $0<\gamma<1$, $t$ is the value from which quadratic shortfalls are measured in the optimal solution. The default probability is $\gamma = 5\%$. Minimizing EQS could be incorporated into a convex problem as a second-order cone constraints, and PortfolioAnalytics uses ECOS in CVXR as the default solver for Second-Order Cone Optimization(SOCopt).

## 7.1 Portfolio Object
The default probability is $\gamma = 5\%$. Specified probability could be given by `arguments`.
```{r}
pspec_eqs <- portfolio.spec(assets=fund_edhec)
pspec_eqs <- add.constraint(pspec_eqs, type="full_investment")
# Add objective of minimizing EQS
pspec_eqs <- add.objective(portfolio = pspec_eqs, type = "risk", name = "EQS",
                          arguments = list(p=0.05))
```

## 7.2 Optimization
```{r}
opt_eqs <- optimize.portfolio(ret_edhec, pspec_eqs, optimize_method = "CVXR")
opt_eqs
```

# 8 Maximizing Mean Return Per Unit Risk
There are three basic types of risk measures: variance or standard deviation, ES and EQS. The problem of maximizing mean return per unit risk can be solved in a clever way by minimizing risk with a target return constraint, as is described below. For all three of these types of problems, both return and risk objectives should be used in PortfolioAnalytics. Then for each of these three optimization problems an appropriate argument needs to be given to the `optimize.portfolio` to specify the type of problem.

## 8.1 Maximum Sharpe Ratio Portfolios
The Sharpe Ratio of a random return $r_P$ of a portfolio $P$ is defined as:
$$\frac{E(r_P) - r_f}{\sqrt{Var(r_P)}}_.$$
The problem of maximizing the Sharpe Ratio can be formulated as a quadratic problem with a budget normalization constraint. It is shown in @cornuejols2018optimization, that this optimization problem is:
\begin{equation*}
\begin{aligned}
\mathop{minimize}\limits_{w} \quad w'\Sigma w\\
s.t. \quad (\hat{\mu} - r_f\textbf{1})^Tw &= 1\\
\textbf{1}^Tw &= \kappa\\
\kappa &> 0
\end{aligned}
\end{equation*}
which has a solution$(w^*,\kappa^*)$ with $k^* \ne 0$, and the maximized Sharpe ratio given by $\tilde{w}^* = w^*/\kappa^*$.

When creating the portfolio, the argument `maxSR = TRUE` should be specified in the function `optimize.portfolio` to distinguish from the mean-variance optimization. NOTE: The default argument is `maxSR = FALSE` since the default action for dealing with both mean and var/StdDev objectives is to maximize quadratic utility.
```{r}
# Create portfolio object
pspec_sr <- portfolio.spec(assets=fund_edhec)
## Add constraints of maximizing Sharpe Ratio
pspec_sr <- add.constraint(pspec_sr, type="full_investment")
pspec_sr <- add.constraint(pspec_sr, type="long_only")
## Add objectives of maximizing Sharpe Ratio
pspec_sr <- add.objective(pspec_sr, type = "return", name = "mean")
pspec_sr <- add.objective(pspec_sr, type="risk", name="var")

# Optimization
optimize.portfolio(ret_edhec, pspec_sr, optimize_method = "CVXR", maxSR=TRUE)
```

## 8.2 Maximum ES ratio Portfolios
The ES ratio(ESratio), which is also called STARR in PortfolioAnalytics, is defined as:
$$\frac{E(r_P) - r_f}{ES_{\gamma}(r_P)}$$
Similar to maximizing Sharpe Ratio, the problem maximizing the ES ratio can be formulated as a minimizing ES problem with a budget normalization constraint.

When creating the portfolio, both return and ES objectives should be given. The default $\gamma=0.05$, and it can be specified by `arguments`. When solving the problem, the default argument `ESratio=TRUE` in the function `optimize.portfolio` specifies the problem type. We note that this argument is equivalent to `maxSTARR=TRUE`, which is used in other vignettes. If one of these two arguments is specified as FALSE, the action will be to minimize ES ignoring the return objective.
```{r}
# Create portfolio object
pspec_ESratio <- portfolio.spec(assets=fund_edhec)
## Add constraints of maximizing return per unit ES
pspec_ESratio <- add.constraint(pspec_ESratio, type="full_investment")
pspec_ESratio <- add.constraint(pspec_ESratio, type="long_only")
## Add objectives of maximizing return per unit ES
pspec_ESratio <- add.objective(pspec_ESratio, type = "return", name = "mean")
pspec_ESratio <- add.objective(pspec_ESratio, type="risk", name="ES",
                               arguments = list(p=0.05))

# Optimization
optimize.portfolio(ret_edhec, pspec_ESratio, optimize_method = "CVXR", ESratio=TRUE)
```

## 8.3 Maximum EQS ratio Portfolios
The EQS ratio of a random return $r_P$ of a portfolio $P$ is defined as:
$$\frac{E(r_P) - r_f}{EQS_{\gamma}(r_P)}$$
Similar to maximizing Sharpe Ratio, the problem maximizing EQS ratio could be formulated as a minimizing EQS problem with a budget normalization constraint.

When creating the portfolio, both return and EQS objectives should be given. The argument `EQSratio=` is used to specify the problem type and the default value is `EQSratio=TRUE`. If `EQSratio=FALSE`, the action will be to minimize EQS ignoring the return objective. The default $\gamma=0.05$, and it can be specified by `arguments`.
```{r}
# Create portfolio object
pspec_EQSratio <- portfolio.spec(assets=fund_edhec)
## Add constraints of maximizing return per unit EQS
pspec_EQSratio <- add.constraint(pspec_EQSratio, type="full_investment")
pspec_EQSratio <- add.constraint(pspec_EQSratio, type="long_only")
## Add objectives of maximizing return per unit EQS
pspec_EQSratio <- add.objective(pspec_EQSratio, type = "return", name = "mean")
pspec_EQSratio <- add.objective(pspec_EQSratio, type="risk", name="EQS",
                                arguments = list(p=0.05))

# Optimization
optimize.portfolio(ret_edhec, pspec_EQSratio, optimize_method = "CVXR", EQSratio=TRUE)
```

# 9 Performance of Portfolios
CVXR solvers provide the Second-Order Cone Optimization (SOCopt) capability required to minimize EQS problem, and managing EQS is of great significance for building portfolios.

In this section, we use the CRSP data set to generate GMV, ES and EQS portfolios and show their performance by plotting cumulative returns and efficient frontiers. In this process, we would like to show the value of EQS in managing portfolios.

## 9.1 Backtesting with GMV, GMES, GMEQS portfolios
In this example, we use daily return of all the CRSP 30 stocks to generate a comparative backtesting among Global Minimum Variance, Global Minimum ES and Global Minimum EQS portfolio. The strategy is to rebalance the portfolio at the end of each month with a rolling window of 500 days, and the performance of backtesting could be shown as a plot of cumulative returns and a plot of drawdown.
```{r, warning=FALSE}
## Generate GMV, GMES and GMEQS portfolios
pspec_sc <- portfolio.spec(assets=fund_CRSP)
pspec_sc <- add.constraint(pspec_sc, type="full_investment")
pspec_sc <- add.constraint(pspec_sc, type="long_only")

pspec_GMV <- add.objective(pspec_sc, type="risk", name="var")
pspec_GMES <- add.objective(pspec_sc, type="risk", name="ES")
pspec_GMEQS <- add.objective(pspec_sc, type="risk", name="EQS")

## Optimize Portfolio at Monthly Rebalancing and 500-Day Training
bt.GMV <- optimize.portfolio.rebalancing(ret_CRSP, pspec_GMV,
                                           optimize_method="CVXR",
                                           rebalance_on="months",
                                           training_period=30,
                                           rolling_window=500)
bt.ES <- optimize.portfolio.rebalancing(ret_CRSP, pspec_GMES,
                                           optimize_method="CVXR",
                                           rebalance_on="months",
                                           training_period=30,
                                           rolling_window=500)
bt.EQS <- optimize.portfolio.rebalancing(ret_CRSP, pspec_GMEQS,
                                           optimize_method="CVXR",
                                           rebalance_on="months",
                                           training_period=30,
                                           rolling_window=500)

## Extract time series of portfolio weights
wts.GMV = extractWeights(bt.GMV)
wts.GMV <- wts.GMV[complete.cases(wts.GMV),]

wts.ES = extractWeights(bt.ES)
wts.ES <- wts.ES[complete.cases(wts.ES),]

wts.EQS = extractWeights(bt.EQS)
wts.EQS <- wts.EQS[complete.cases(wts.EQS),]

## Compute cumulative returns of three portfolios
GMV = Return.rebalancing(retM_CRSP, wts.GMV)
ES = Return.rebalancing(retM_CRSP, wts.ES)
EQS = Return.rebalancing(retM_CRSP, wts.EQS)

# Combine GMV, ES and EQS portfolio cumulative returns
ret.comb <- na.omit(merge(GMV, ES, EQS, all=F))
names(ret.comb) = c("GMV", "GMES", "GMEQS")

# Compute cumulative gross portfolios returns
R <- ret.comb
gross_cum <- TRUE
c.xts <- if ( gross_cum ) {
  cumprod(1+R)
} else {
  cumsum(R)
}

# Cumulative returns panel (Peter Carl)
p <- xts::plot.xts(c.xts[,1], col="black", main = "Cumulative returns",
                   grid.ticks.lwd=1, grid.ticks.lty = "solid", grid.ticks.on = "years",
                   labels.col="grey20", cex.axis=0.8, format.labels = "%b\n%Y",
                   lty = "dotted", ylim = c(min(c.xts), max(c.xts)))
p <- xts::addSeries(c.xts[,2], on=1, lwd=2, col="dark blue", lty="dashed")
p <- xts::addSeries(c.xts[,3], on=1, lwd=2, col="dark green", lty="solid")
p <- xts::addLegend("topleft", on = 1,
                    legend.names = names(c.xts),
                    lty = c(3, 2, 1), lwd = rep(2, NCOL(c.xts)),
                    col = c("black", "dark blue", "dark green"),
                    bty = "o", box.col = "white",
                    bg=rgb(t(col2rgb("white")), alpha = 200,
                           maxColorValue = 255) )

## Drawdowns panel(Peter Carl)
d.xts <- PerformanceAnalytics::Drawdowns(R)
p <- xts::addSeries(d.xts[,1], col="black", lwd=2, main="Drawdown",
                    ylim = c(min(d.xts), 0), lty=3)
p <- xts::addSeries(d.xts[,2], on=2, lwd=2, col="dark blue", lty=2)
p <- xts::addSeries(d.xts[,3], on=2, lwd=2, col="dark green", lty=1)

## panel 1 and 2 ylim
ylim1 <- c(p$Env$ylim[[2]][1], p$Env$ylim[[2]][2])
ylim2 <- c(p$Env$ylim[[4]][1], p$Env$ylim[[4]][2])
ylim <- c(ylim1, ylim2)
# get longest drawdown dates for xts object
dt <- table.Drawdowns(R, top = 1) # just want to find the worst drawdown
dt2 <- t(dt[,c("From", "To")])
x <- as.vector(dt2[,NCOL(dt2)])
y <- as.xts(matrix(rep(ylim, length(x)),ncol=length(ylim), byrow=TRUE), order.by=as.Date(x))
i=1
p <- xts::addPolygon(y[i:(i+1),1:2], on=-1, col="lightgrey") # top panel
p <- xts::addPolygon(y[i:(i+1),3:4], on=-2, col="lightgrey") # lower panel

p
```
\begin{center}
Fig 9.1
\end{center}

## 9.2 Backtesting with SR, ESratio, EQSratio portfolios
In this example, we use daily return of all the CRSP 30 stocks to generate a comparative backtesting among Maximum Sharpe Ratio, Maximum ES Ratio and Maximum EQS Ratio portfolio. The strategy is to rebalance the portfolio at the end of each month with a rolling window of 500 days, and the performance of backtesting could be shown as a plot of cumulative returns and a plot of drawdown.
```{r, warning=FALSE}
## Generate GMV, GMES and GMEQS portfolios
pspec_sc_ratio <- add.objective(pspec_sc, type="return", name="mean")
pspec_Sr <- add.objective(pspec_sc_ratio, type="risk", name="var")
pspec_ESr <- add.objective(pspec_sc_ratio, type="risk", name="ES")
pspec_EQSr <- add.objective(pspec_sc_ratio, type="risk", name="EQS")

## Optimize Portfolio at Monthly Rebalancing and 500-Day Training
bt.Sr <- optimize.portfolio.rebalancing(ret_CRSP, pspec_Sr, maxSR=TRUE,
                                           optimize_method="CVXR",
                                           rebalance_on="months",
                                           training_period=30,
                                           rolling_window=500)
bt.ESr <- optimize.portfolio.rebalancing(ret_CRSP, pspec_ESr,
                                           optimize_method="CVXR",
                                           rebalance_on="months",
                                           training_period=30,
                                           rolling_window=500)
bt.EQSr <- optimize.portfolio.rebalancing(ret_CRSP, pspec_EQSr,
                                           optimize_method="CVXR",
                                           rebalance_on="months",
                                           training_period=30,
                                           rolling_window=500)

## Extract time series of portfolio weights
wts.Sr = extractWeights(bt.Sr)
wts.Sr <- wts.Sr[complete.cases(wts.Sr),]

wts.ESr = extractWeights(bt.ESr)
wts.ESr <- wts.ESr[complete.cases(wts.ESr),]

wts.EQSr = extractWeights(bt.EQSr)
wts.EQSr <- wts.EQSr[complete.cases(wts.EQSr),]

## Compute cumulative returns of three portfolios
Sr = Return.rebalancing(retM_CRSP, wts.Sr)
ESr = Return.rebalancing(retM_CRSP, wts.ESr)
EQSr = Return.rebalancing(retM_CRSP, wts.EQSr)

# Combine Sr, ESr and EQSr portfolio cumulative returns
ret.comb <- na.omit(merge(Sr, ESr, EQSr, all=F))
names(ret.comb) = c("Sharpe ratio", "ES ratio", "EQS ratio")

# Compute cumulative gross portfolios returns
R <- ret.comb
gross_cum <- TRUE
c.xts <- if ( gross_cum ) {
  cumprod(1+R)
} else {
  cumsum(R)
}

# Cumulative returns panel (Peter Carl)
p <- xts::plot.xts(c.xts[,1], col="black", main = "Cumulative returns",
                   grid.ticks.lwd=1, grid.ticks.lty = "solid", grid.ticks.on = "years",
                   labels.col="grey20", cex.axis=0.8, format.labels = "%b\n%Y",
                   lty = "dotted", ylim = c(min(c.xts), max(c.xts)))
p <- xts::addSeries(c.xts[,2], on=1, lwd=2, col="dark blue", lty="dashed")
p <- xts::addSeries(c.xts[,3], on=1, lwd=2, col="dark green", lty="solid")
p <- xts::addLegend("topleft", on = 1,
                    legend.names = names(c.xts),
                    lty = c(3, 2, 1), lwd = rep(2, NCOL(c.xts)),
                    col = c("black", "dark blue", "dark green"),
                    bty = "o", box.col = "white",
                    bg=rgb(t(col2rgb("white")), alpha = 200,
                           maxColorValue = 255) )

## Drawdowns panel(Peter Carl)
d.xts <- PerformanceAnalytics::Drawdowns(R)
p <- xts::addSeries(d.xts[,1], col="black", lwd=2, main="Drawdown",
                    ylim = c(min(d.xts), 0), lty=3)
p <- xts::addSeries(d.xts[,2], on=2, lwd=2, col="dark blue", lty=2)
p <- xts::addSeries(d.xts[,3], on=2, lwd=2, col="dark green", lty=1)

## panel 1 and 2 ylim
ylim1 <- c(p$Env$ylim[[2]][1], p$Env$ylim[[2]][2])
ylim2 <- c(p$Env$ylim[[4]][1], p$Env$ylim[[4]][2])
ylim <- c(ylim1, ylim2)
# get longest drawdown dates for xts object
dt <- table.Drawdowns(R, top = 1) # just want to find the worst drawdown
dt2 <- t(dt[,c("From", "To")])
x <- as.vector(dt2[,NCOL(dt2)])
y <- as.xts(matrix(rep(ylim, length(x)),ncol=length(ylim), byrow=TRUE), order.by=as.Date(x))
i=1
p <- xts::addPolygon(y[i:(i+1),1:2], on=-1, col="lightgrey") # top panel
p <- xts::addPolygon(y[i:(i+1),3:4], on=-2, col="lightgrey") # lower panel

p
```
\begin{center}
Fig 9.2
\end{center}

## 9.3 Efficient Frontier
We generate efficient frontiers with mean-StdDev, mean-ES and mean-EQS portfolios by using 30 stocks from CRSP data set. Considering that the data may show different properties over a long period of time, we only use the monthly return in the last 5 years to generate efficient frontiers, that is from 2011-01 to 2015-12 and defined in Section 2.3 as `retM_CRSP_5`. We can use `create.EfficientFrontier` to calculate the mean value and risk value for the frontier, then use `chart.EfficientFrontier` to draw the frontier.

### 9.3.1 Mean-StdDev Efficient Frontier
```{r}
# mean-var efficient frontier
meanvar.ef <- create.EfficientFrontier(R=retM_CRSP_5, portfolio=pspec_sc, type="mean-StdDev")
meanvar.ef
chart.EfficientFrontier(meanvar.ef, match.col="StdDev", type="l",
                        chart.assets = FALSE, main="Mean-StdDev Efficient Frontier",
                        RAR.text="Sharpe ratio", pch=1)
```
\begin{center}
Fig 9.3
\end{center} 
The Sharpe ratio could be calculated by the frontier value and the maximum Sharpe ratio could be found.
```{r}
meanvar.ef$frontier[, 1:2]
sr = meanvar.ef$frontier[, 1]/meanvar.ef$frontier[, 2]
cat("maximum Sharpe ratio:", max(sr))
cat("mean of the maximum SR portfolio:", meanvar.ef$frontier[, 1][sr == max(sr)])
cat("StdDev of the maximum SR portfolio:", meanvar.ef$frontier[, 2][sr == max(sr)])
```

Note that we have introduced the method of finding the theoretical maximum Sharpe ratio portfolio in Section 8.1, which may be a little different from the estimated maximum Sharpe ratio calculated by the discrete efficient frontier value. It is because the function of efficient frontier uses the mean value of the maximum mean return portfolio and the minimum variance portfolio as boundary values, then divides the mean interval equally and calculates the corresponding StdDev value, and then gives discrete mean-StdDev points to fit the efficient frontier curve. The "maximum" Sharpe ratio found by the efficient frontier function is the maximum value calculated by limited number of discrete points. The default number of points is 25, and the specific number could be given by `n.portfolios = {number}`.

We can identify the maximum Sharpe ratio portfolio in blue point on the mean-StdDev efficient frontier.
```{r, warning=FALSE}
# Mean-StdDev Efficient Frontier
pspec_MV <- add.objective(pspec_sc, type="risk", name="var")
pspec_MV <- add.objective(portfolio=pspec_MV, type="return", name="mean")
opt_MV <- optimize.portfolio(retM_CRSP_5, pspec_MV, optimize_method = "CVXR", 
                              maxSR=TRUE, trace = TRUE)
opt_MV
chart.EfficientFrontier(opt_MV, match.col="StdDev", chart.assets = FALSE, 
                        main="Mean-StdDev Efficient Frontier",
                        RAR.text="Sharpe Ratio", pch=1, xlim = c(0, 0.1))
```
\begin{center}
Fig 9.4
\end{center}
The theoretical maximum Sharpe ratio portfolio is very close to the result generated by the efficient frontier, and the Sharpe ratio value is almost the same but the mean and StdDev value are slightly different.

With different constraint types, we can create mean-StdDev efficient frontiers for multiple portfolios and overlay the plots.
```{r}
pspec_sc_init <- portfolio.spec(assets=fund_CRSP)
pspec_sc_init <- add.constraint(pspec_sc_init, type="full_investment")

# Portfolio with long-only constraints
pspec_sc_lo <- add.constraint(portfolio=pspec_sc_init, type="long_only")

# Portfolio with long-only box constraints
pspec_sc_lobox <- add.constraint(portfolio=pspec_sc_init, type="box", min=0.02, max=0.1)

# Portfolio with long-short box constraints
pspec_sc_lsbox <- add.constraint(portfolio=pspec_sc_init, type="box", min=-0.1, max=0.1)

# Combine the portfolios into a list
portf_list <- combine.portfolios(list(pspec_sc_lo, pspec_sc_lobox, pspec_sc_lsbox))

# Plot the efficient frontier overlay of the portfolios with varying constraints
legend_labels <- c("Long Only", "Long Only Box", "Long Short Box")
chart.EfficientFrontierOverlay(R=retM_CRSP_5, portfolio_list=portf_list, 
                               type="mean-StdDev", match.col="StdDev", 
                               legend.loc="bottomright", chart.assets = FALSE,
                               legend.labels=legend_labels, cex.legend=1,
                               labels.assets=FALSE, lwd = c(3,3,3),
                               col = c("black", "dark red", "dark green"),
                               main="Overlay Mean-StdDev Efficient Frontiers",
                               xlim = c(0.03, 0.11), ylim = c(0.005, 0.035))
```
\begin{center}
Fig 9.5
\end{center}
The plot clearly show that the portfolio under the long-short box constraints has the best performance, though it also requires shorting which may not be possible for many real-world portfolios.

### 9.3.2 Mean-ES Efficient Frontier
Generate the mean-ES efficient frontier:
```{r, warning=FALSE}
# Mean-ES Efficient Frontier
meanetl.ef <- create.EfficientFrontier(R=retM_CRSP_5, portfolio=pspec_sc, type="mean-ES")
chart.EfficientFrontier(meanetl.ef, match.col="ES", type="l",
                        chart.assets = FALSE, main="Mean-ES Efficient Frontier",
                        RAR.text="ES ratio", pch=1)
```
\begin{center}
Fig 9.6
\end{center}
Generate multiple mean-ES efficient frontiers and overlay the plots.
```{r}
legend_labels <- c("Long Only ES (p=0.05)", 
                   "Long Only Box ES (p=0.05)", "Long Short Box ES (p=0.05)")
chart.EfficientFrontierOverlay(R=retM_CRSP_5, portfolio_list=portf_list, 
                               type="mean-ES", match.col="ES",
                               legend.loc="bottomright", chart.assets = FALSE,
                               legend.labels=legend_labels, cex.legend=1,
                               labels.assets=FALSE, lwd = c(3,3,3),
                               col = c("black", "dark red", "dark green"),
                               main="Overlay Mean-ES Efficient Frontiers",
                               xlim = c(0.03, 0.17), ylim = c(0.005, 0.035))
```
\begin{center}
Fig 9.7
\end{center}
For each mean-ES efficient frontier, the left endpoint is the global minimum ES and its corresponding mean return, the right endpoint is the maximum mean return and and its corresponding ES, that is, the minimum ES when the maximum mean return is the target return.

With the same target return, the portfolio under the long-short box constraints can have a minimum ES value. It performs best, but shorting may not be possible for many real-world portfolios. The mean-ES efficient frontier of the long-only portfolio covers the widest range, that is because it can reach a larger return range than the long-box portfolio, but it is not as easy to reach the target return as the long-short portfolio, it needs to take more risks.

Instead of generating efficient frontiers with different constraint types, we can also generate mean-ES efficient frontiers with different tail probability $\gamma$.
```{r}
# Create long-only ES portfolios with different tail probabilities
ES_05 <- add.objective(portfolio=pspec_sc_lo, type="risk", name="ES", 
                          arguments=list(p=0.05))

ES_10 <- add.objective(portfolio=pspec_sc_lo, type="risk", name="ES", 
                          arguments=list(p=0.1))

ES_15 <- add.objective(portfolio=pspec_sc_lo, type="risk", name="ES", 
                      arguments=list(p=0.15))

# Combine the portfolios into a list
portf_ES_list <- combine.portfolios(list(ES_05, ES_10, ES_15))

# Plot the efficient frontier overlay of the portfolios with varying tail probabilities
legend_ES_labels <- c("ES (p=0.05)", "ES (p=0.1)", "ES (p=0.15)")
chart.EfficientFrontierOverlay(R=retM_CRSP_5, portfolio_list=portf_ES_list, 
                               type="mean-ES", match.col="ES", 
                               legend.loc="bottomright", chart.assets = FALSE,
                               legend.labels=legend_ES_labels, cex.legend=1,
                               labels.assets=FALSE, lwd = c(3,3,3),
                               col = c("black", "dark red", "dark green"),
                               main="Overlay Mean-ES Efficient Frontiers",
                               xlim = c(0.035, 0.165), ylim = c(0.005, 0.03))
```
\begin{center}
Fig 9.8
\end{center}
ES portfolio with a larger tail probability will have better performance, but this is only because larger tail probability gives a wider ES calculation range, so the expected ES value is smaller. Actually, these three lines cannot be compared because their x-lables are ES with different tail probabilities.

This plot mainly shows that the same portfolio will have different ES values under different tail probabilities. For example, the portfolio that reaches the maximum return, which is the right endpoint of each mean-ES efficient frontier. They are the same portfolio, but the ES values are different. It informs us the importance of tail probability for calculating ES. Different tail probabilities only focus on risks in different degrees, rather than changes in risk levels.

### 9.3.3 Mean-EQS Efficient Frontier
```{r, warning=FALSE}
# Mean-EQS Efficient Frontier
meaneqs.ef <- create.EfficientFrontier(R=retM_CRSP_5, portfolio=pspec_sc, type="mean-EQS")
chart.EfficientFrontier(meaneqs.ef, match.col="EQS", type="l",
                        chart.assets = FALSE, main="Mean-EQS Efficient Frontier",
                        RAR.text="EQS ratio", pch=1)
```
\begin{center}
Fig 9.9
\end{center}
Mean-EQS efficient frontier is more like a piecewise function rather than a smooth curve.

### 9.3.4 Efficient Frontiers Comparison among minStd, minES and minEQS portfolios
In order to calculate the value of StdDev, ES and EQS of a portfolio, we provide a new function `extract_risk(R =, w =)`. Asset returns `R` and the portfolio weights `w` should be specified when using the function, `ES_alpha` and `EQS_alpha` are alternative arguments and the default values are 0.05.
```{r}
# usage example: minStd Portfolio
minstd_port <- add.objective(pspec_sc, type = "risk", name = "StdDev")
minstd_w <- optimize.portfolio(retM_CRSP_5, minstd_port, optimize_method = "CVXR")$weight

# risk values with default alpha=0.05
extract_risk(retM_CRSP_5, minstd_w)

# risk values with specific alpha
extract_risk(retM_CRSP_5, minstd_w, ES_alpha = 0.1, EQS_alpha = 0.1)
```

For comparing a specific risk type in minStd, minES and minEQS portfolios and generating multiple efficient frontiers, we provide a new function

`chart.EfficientFrontierCompare(R =, portfolio =, risk_type =, match.col =, guideline =)`

where `risk_type` is the risk to be compared, and `match.col` is the vector of the portfolios that participate in the comparison. If there are only two frontiers in the comparison, the default of the argument `guideline` is `TRUE`.
```{r}
# example 1: Compare StdDev of minStd and minES portfolios with guideline
chart.EfficientFrontierCompare(R=retM_CRSP_5, portfolio=pspec_sc, risk_type="StdDev",
                               match.col = c("StdDev", "ES"), lwd = c(2, 2))

# example 2: Compare ES of minStd, minES and minEQS portfolios without guideline
chart.EfficientFrontierCompare(R=retM_CRSP_5, portfolio=pspec_sc, risk_type="ES",
                               match.col = c("StdDev", "ES", "EQS"), guideline = FALSE,
                               col = c(1,2,4), lty = c(1, 2, 4), lwd = c(2, 2, 2))
```

\newpage
# Reference